---
title: "Multimodal Models and Numerosity"
date: "September 12, 2024"
output:
  # pdf_document: 
  #    fig_caption: yes
  #    keep_md: yes
  #    keep_tex: yes
  html_document:
     keep_md: yes
     toc: yes
     toc_float: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi = 300, fig.format = "pdf")
```

```{r include=FALSE}
library(tidyverse)
library(lmtest)
library(forcats)
library(broom)
library(lme4)
library(ggridges)
library(lmerTest)
library(broom.mixed)
library(ggrepel)
library(tools)
library(viridis)

library(ggcorrplot)

all_colors <- viridis::viridis(10, option = "mako")
my_colors <- all_colors[c(3, 5, 7)]  # Selecting specific colors from the palette
```


# Load data

```{r}
# setwd("/Users/seantrott/Dropbox/UCSD/Research/NLMs/vlm-vit-num/analysis")
directory_path <- "../results"
csv_files <- list.files(path = directory_path, pattern = "*.csv", full.names = TRUE)
csv_list <- csv_files %>%
  map(~ read_csv(.))
df_hf_models <- bind_rows(csv_list)


df_hf_models = df_hf_models %>%
  mutate(numerosity_diff = abs(numerosity_2 - numerosity_1)) %>%
  mutate(log_params = log10(n_params)) %>%
  group_by(model_name) %>%
  mutate(max_layer = max(layer)) %>%
  mutate(model_type = case_when(
    str_detect(model_name, "clip") == TRUE ~ "VLM",
    TRUE ~ "ViT"))  %>%
  mutate(cosine_similarity_z = scale(cosine_similarity),
         numerosity_diff_z = scale(numerosity_diff),
         area_diff_z = scale(area_diff)) %>%
  mutate(model_name2 = str_extract(model_name, "^(?:[^-]*-?){1,3}[^-]*"))


table(df_hf_models$model_name)
table(df_hf_models$model_name, df_hf_models$numerosity_comparison_type)
table(df_hf_models$model_name, df_hf_models$image_type)

```


# Descriptive analyses

```{r}
df_hf_models %>%
  filter(layer == max_layer) %>%
  ggplot(aes(x = cosine_similarity)) +
  geom_histogram(alpha = .6) +
  theme_minimal() +
  labs(x = "Cosine Similarity") +
  theme(text = element_text(size = 15)) +
  facet_wrap(~model_name)

```

# Analyses

## Same Vs. Different Numerosity

```{r}
df_summary <- df_hf_models %>%
  group_by(model_name2, numerosity_comparison_type, layer, max_layer, image_type) %>%
  summarize(avg_similarity = mean(cosine_similarity, na.rm = TRUE),
            se_similarity = sd(cosine_similarity, na.rm = TRUE) / sqrt(n()))

df_summary %>%
  filter(layer == max_layer) %>%
  ggplot(aes(x = model_name2,
             y = avg_similarity,
             fill = numerosity_comparison_type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.5), width = .6) +  
  # geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = avg_similarity - se_similarity, 
                    ymax = avg_similarity + se_similarity), 
                width = 0.2,
                position = position_dodge(width = 0.5)) + 
  labs(# title = "",
       x = "Model",
       y = "Average Cosine Similarity",
       fill = "",
       color = "") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis(discrete=TRUE) +
  # scale_color_viridis_d() +
  scale_y_continuous(n.breaks = 3) +
  theme(axis.title = element_text(size=rel(1.2)),
        axis.text = element_text(size = rel(1.2)),
        legend.text = element_text(size = rel(1.2)),
        # legend.title = element_text(size = rel(1.5)),
        strip.text.x = element_text(size = rel(1.2)),
        legend.position = "bottom") +
  facet_wrap(~image_type)


df_hf_models %>%
  filter(layer == max_layer) %>%
  ggplot(aes(x = cosine_similarity,
             y = model_name,
             fill = numerosity_comparison_type)) +
  geom_density_ridges2(aes(height = ..density..), 
                       color=gray(0.25), 
                       alpha = .7, 
                       scale=.85, 
                       # size=1, 
                       size = 0,
                       stat="density") +
  labs(x = "Cosine Similarity",
       y = "",
       fill = "Comparison Type") +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE) +
  theme(text = element_text(size = 15),
        legend.position="bottom") +
  facet_wrap(~image_type)

### Overall sensitivity
m1 = lmer(data = df_hf_models,
          cosine_similarity ~ area_diff + numerosity_comparison_type + 
            (1 | image_1) + (1 | image_2) + 
            (1 |model_name) + (1|image_type),
          control=lmerControl(optimizer="bobyqa"))

summary(m1)


### Interaction with layer
m1 = lmer(data = df_hf_models,
          cosine_similarity ~ area_diff *layer + numerosity_comparison_type * layer + 
            (1 | image_1) + (1 | image_2) + 
            (1 |model_name) + (1|image_type))

summary(m1)


## Plot differences by layer
summary_df <- df_hf_models %>%
  group_by(model_name, layer, numerosity_comparison_type) %>%
  summarize(mean_cosine = mean(cosine_similarity, na.rm = TRUE), .groups = 'drop')

difference_df <- summary_df %>%
  pivot_wider(names_from = numerosity_comparison_type, values_from = mean_cosine) %>%
  mutate(difference = same - different) %>%
  group_by(model_name) %>%
  mutate(max_layer = max(layer),
         prop_layer = layer / max_layer) %>%
  mutate(binned_prop_layer = ntile(prop_layer, 10)) %>%
  mutate(prop_binned = binned_prop_layer / 10) 


smoothed_data <- difference_df %>%
  group_by(prop_binned) %>%
  summarize(avg_difference = mean(difference, na.rm = TRUE)) %>%
  mutate(model_name = "Average")

# Create the plot
difference_df %>%
  ggplot(aes(x = prop_binned, y = difference, 
             color = model_name, group = model_name)) +
  geom_line(alpha = 0.5) + # Individual lines for each model_name
  geom_line(data = smoothed_data, aes(x = prop_binned, y = avg_difference), 
             color = "black", size = 1.2) + # Smoothed average
  labs(
    title = "",
    x = "Layer Depth Ratio",
    y = "Difference (Same - Different)",
    color = ""
  ) +
  theme_minimal() +
  scale_color_viridis(option = "mako", discrete=TRUE)  +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.position = "none")



```


## Continuous function of numerosity


```{r}
df_hf_models <- df_hf_models %>%
  mutate(model_name2 = str_extract(model_name, "^(?:[^-]*-?){1,3}[^-]*"))

df_hf_models %>%
  filter(layer == max_layer) %>%
  ggplot(aes(x = numerosity_diff,
             y = cosine_similarity,
             color = model_type)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm") +
  theme_minimal() +
  scale_color_viridis(option = "mako", discrete=TRUE) +
  labs(# title = "",
       x = "Numerosity Difference",
       y = "Cosine Similarity",
       fill = "",
       color = "Model Type") +
  theme(text = element_text(size = 12),
        legend.position = "bottom") +
  facet_wrap(~reorder(model_name2, n_params))



### Overall sensitivity
m1 = lmer(data = df_hf_models,
          cosine_similarity ~ area_diff + numerosity_diff + 
            (1 | image_1) + (1 | image_2) + 
            (1 |model_name) + (1|image_type),
          control=lmerControl(optimizer="bobyqa"))

summary(m1)


### Overall sensitivity
m1 = lmer(data = filter(df_hf_models, layer == max_layer),
          cosine_similarity ~ area_diff + numerosity_diff + 
            (1 | image_1) + (1 | image_2) + 
            (1 |model_name) + (1|image_type),
          control=lmerControl(optimizer="bobyqa"))

summary(m1)


### Interaction
m1 = lmer(data = df_hf_models,
          cosine_similarity ~ area_diff * layer +  numerosity_diff * layer +
            (1 | image_1) + (1 | image_2) + 
            (1 |model_name) + (1|image_type),
          control=lmerControl(optimizer="bobyqa"))

summary(m1)




```


# VLM vs. ViT

## Same vs. Different

```{r}

df_hf_models = df_hf_models %>%
  mutate(model_type = case_when(
    str_detect(model_name, "clip") == TRUE ~ "VLM",
    TRUE ~ "ViT")) %>%
  group_by(model_name) %>%
  mutate(max_layer = max(layer))

df_summary <- df_hf_models %>%
  filter(layer == max_layer) %>%
  group_by(model_type, numerosity_comparison_type) %>%
  summarize(avg_similarity = mean(cosine_similarity, na.rm = TRUE),
            se_similarity = sd(cosine_similarity, na.rm = TRUE) / sqrt(n()))

df_summary %>%
  ggplot(aes(x = factor(model_type),
             y = avg_similarity,
             color = numerosity_comparison_type)) +
  geom_point(position = position_dodge(width = 0.5), size = 2) +  
  geom_errorbar(aes(ymin = avg_similarity - 2 * se_similarity, 
                    ymax = avg_similarity + 2 * se_similarity), 
                width = 0.2,
                position = position_dodge(width = 0.5)) + 
  labs(# title = "",
       x = "Model Type",
       y = "Average Cosine Similarity",
       color = "") +
  theme_minimal() +
  scale_color_viridis(discrete=TRUE) +
  theme(axis.title = element_text(size=rel(1.2)),
        axis.text = element_text(size = rel(1.2)),
        legend.text = element_text(size = rel(1.2)),
        # legend.title = element_text(size = rel(1.5)),
        strip.text.x = element_text(size = rel(1.2)),
        legend.position = "bottom") 




m1 = lmer(data = filter(df_hf_models, layer == max_layer),
          cosine_similarity ~ area_diff * model_type + 
            numerosity_comparison_type * model_type + 
            patch_size * numerosity_comparison_type  + 
            log_params * numerosity_comparison_type + 
            (1 | image_1) + (1 | image_2) + (1|image_type) +
            (1|model_name),
          control=lmerControl(optimizer="bobyqa"))

summary(m1)
```

### Multiverse analysis

Significant interaction does not always hold up under alternative model specifications, though it is generally trending / positive.

```{r}
m1 = lmer(data = filter(df_hf_models, layer == max_layer),
          cosine_similarity ~ # area_diff * model_type + 
            numerosity_comparison_type * model_type + 
            patch_size * numerosity_comparison_type  + 
            log_params * numerosity_comparison_type +
            (1 | image_1) + (1 | image_2) + # (1|image_type) +
            (1|model_name))

summary(m1)

m1 = lmer(data = filter(df_hf_models, layer == max_layer),
          cosine_similarity ~ area_diff * model_type + 
            numerosity_comparison_type * model_type + 
            # patch_size * numerosity_comparison_type  + 
            log_params * numerosity_comparison_type +
            (1 | image_1) + (1 | image_2) + # (1|image_type) + 
            (1|model_name))

summary(m1)

m1 = lmer(data = filter(df_hf_models, layer == max_layer),
          cosine_similarity ~ # area_diff * model_type + 
            numerosity_comparison_type * model_type + 
            # patch_size * numerosity_comparison_type  +
            log_params * numerosity_comparison_type +
            (1 | image_1) + (1 | image_2) + # (1|image_type) + 
            (1|model_name))

summary(m1)

m1 = lmer(data = filter(df_hf_models, layer == max_layer),
          cosine_similarity ~ area_diff * model_type + 
            numerosity_comparison_type * model_type + 
            patch_size * numerosity_comparison_type  +
            # log_params * numerosity_comparison_type +
            (1 | image_1) + (1 | image_2) + # (1|image_type) +
            (1|model_name))

summary(m1)
```


## Continuous

```{r}

df_hf_models %>%
  filter(layer == max_layer) %>%
  ggplot(aes(x = numerosity_diff,
             y = cosine_similarity,
             color = model_type)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(x = "Numerosity Difference",
       y = "Cosine Similarity",
       color = "") +
  scale_color_viridis(option = "mako", discrete=TRUE) +
  theme(text = element_text(size = 15),
        legend.position = "bottom") +
  facet_wrap(~model_type)



df_hf_models_last = df_hf_models %>%
  filter(layer == max_layer) 

m1 = lmer(data = df_hf_models_last,
          cosine_similarity ~ area_diff * model_type + numerosity_diff * model_type + 
            patch_size * numerosity_diff + log_params * numerosity_diff +
            (1 | image_1) + (1 | image_2) + (1|image_type) + 
            (1|model_name))

summary(m1)



### For visualizing, remove key factors from model 
m_reduced = lmer(data = df_hf_models_last,
          cosine_similarity ~ area_diff + # numerosity_diff * model_type + 
            patch_size + # log_params * numerosity_diff +
            (1 | image_1) + (1 | image_2) + (1|image_type) + 
            (1|model_name))

df_hf_models_last$resid = residuals(m_reduced)

df_hf_models_last %>%
  ggplot(aes(x = numerosity_diff,
             y = resid,
             color = model_type)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(x = "Numerosity Difference",
       y = "Residuals",
       color = "") +
  scale_color_viridis(option = "mako", discrete=TRUE) +
  theme(text = element_text(size = 15),
        legend.position = "bottom") +
  facet_wrap(~model_type)

```


### Multiverse analysis

Significant interaction is preserved under alternative specifications.

```{r}
m1 = lmer(data = filter(df_hf_models, layer == max_layer),
          cosine_similarity ~ # area_diff * model_type + 
            numerosity_diff * model_type + 
            patch_size * numerosity_diff + 
            log_params * numerosity_diff +
            (1 | image_1) + (1 | image_2) + # (1|image_type) + 
            (1|model_name))

summary(m1)

m1 = lmer(data = filter(df_hf_models, layer == max_layer),
          cosine_similarity ~ area_diff * model_type + 
            numerosity_diff * model_type + 
            # patch_size * numerosity_diff + 
            log_params * numerosity_diff +
            (1 | image_1) + (1 | image_2) + # (1|image_type) + 
            (1|model_name))

summary(m1)

m1 = lmer(data = filter(df_hf_models, layer == max_layer),
          cosine_similarity ~ # area_diff * model_type + 
            numerosity_diff * model_type + 
            # patch_size * numerosity_diff + 
            log_params * numerosity_diff +
            (1 | image_1) + (1 | image_2) + # (1|image_type) + 
            (1|model_name))

summary(m1)

m1 = lmer(data = filter(df_hf_models, layer == max_layer),
          cosine_similarity ~ area_diff * model_type + 
            numerosity_diff * model_type + 
             patch_size * numerosity_diff + 
            # log_params * numerosity_diff +
            (1 | image_1) + (1 | image_2) + # (1|image_type) + 
            (1|model_name))

summary(m1)
```


